---
title: Group Assignment MidTerm - California Small Business - Credit Card Accounts
  Data
author: "C. Loya, S.F. Ali, R. Lakshminarayana & A. Carriedo"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

# Clear the work space 
```{r cleanup}
rm(list = ls())
```
# Load Libraries and Set Options  

Here are the libraries used in this program. 

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
library(readxl) # read from excel
library(writexl) # write to excel
library(ggplot2) # make ncie plots 
library(xts) # helpful for making dates for time series as.yearmon(). 
library(forecast) # This has the ARMA function 
library(psych) # nice summary statistics with describe() function
library(dplyr) # program to filter and pipe the data in the GGPLOT for the forecasts
library(tinytex)
uninstall_tinytex()
library(urca) #  for the unit root test function ur.df
```

```{r Read the data, echo=TRUE }
data_from_excel <- read_excel("WF_Raw_data.xlsx") 
```

## R Program Summary
\newline

1. A general description of the data.

2. A plot of the series.

3. Summary statistics.

4. Any problems with the series. Are there any outliers or unusual events?

5. Explanations of any simplifying assumptions used in modeling the series. (How are outliers treated?

6. One set of forecasts and confidence intervals for the series from a model presented in the first three weeks of the course.

A. An appendix to include technical details of the analysis that supports your business document.


### 1. A general description of the data.  

The data provided here is the monthly number of new Credit Card accounts openend by small businesses in California with a large Banking institue, between year 2002 and 2020. 

```{r creat_the_structure_of_the_data.frame, message=FALSE, warning=FALSE, tidy=TRUE}
str(data_from_excel)



CCSales_ts <- ts(data_from_excel, start=c(2002,1), frequency = 12, names =c("CA_New_Credit_Card_Accounts") )
str(CCSales_ts)
```

### 2. A plot of the series.

```{r creat timeseries and plot, echo=TRUE, warning=FALSE}

autoplot(CCSales_ts, colour = 'red') +
  labs(x = "Observation Date", y = "CA_New_Credit_Card_Accounts")

#
#  Plot with ggplot2
#

df_data <- data.frame(date=as.Date(as.yearmon(time(CCSales_ts))), Y=as.matrix(CCSales_ts))


ggplotdf <- ggplot(data=df_data[1:2], aes(x=date,y=CA_New_Credit_Card_Accounts))+geom_line()


ggplotdf + scale_x_date(date_label = "%Y", date_breaks = "1 year") + theme(axis.text.x = element_text( angle = 45))
```

### 3. Summary statistics.
```{r summary of data}
summary(CCSales_ts)
```

### 4. Any problems with the series. Are there any outliers or unusual events?
```{r}
hist(data_from_excel$`New Visa Card Opens`)
boxplot(data_from_excel,horizontal = T)
```


Based on 1.5*IQR rule, there are no Outliers in this dataset.

### 5. Explanations of any simplifying assumptions used in modeling the series.


In the ggplot above, presence of seasonlity in the dataset is clearly evident. But, a global trend is not observed. hence, Holt Exponential Smoothing ( with lower value of beta) will be utilized for forecasting.

### 6. Forecasts and confidence intervals for the series from Holt Exponential Smoothing model

```{r Holt Exponential Smoothing for Forecast}
Holt_Winters <- ets(CCSales_ts, model = "MAM", alpha = .4, beta = .15, gamma=.3)  
Holt_Winters
plot(Holt_Winters)

HOLT <- ets(CCSales_ts, model = "AAN", alpha = .2, beta = .05)  
HOLT
plot(HOLT)
HOLT.pred <- forecast(HOLT, h = 18, level = c(95))
plot(HOLT.pred)

HOLT_auto <- ets(CCSales_ts, model = "AAN")  
HOLT_auto
plot(HOLT_auto)
HOLT_auto.pred <- forecast(HOLT_auto, h = 3, level = c(95))
plot(HOLT_auto.pred)

col_names <- as.yearmon(time(HOLT.pred$mean)) 
df <-  data.frame( as.Date(col_names), HOLT.pred$mean, HOLT.pred$lower, HOLT.pred$upper)
colnames(df) <- c("Date", "Forecast", "95% Lower", "95% Upper" )
# write_xlsx(df, "Annual_New_CredirCard_prediction.xlsx")

df

ggplot(data = df, aes(x = Date, y = as.double(Forecast) )) +
  geom_point( color="blue" ) + 
  geom_line(aes(x = Date, y = `95% Lower`), color="red" ) +  
  geom_line(aes(x = Date, y = `95% Upper`), color="red" ) +
  geom_point(data=df_data, mapping=aes(x=date, y=CA_New_Credit_Card_Accounts))+
  labs(x = "Date", y = "CA_New_Credit_Card_Accounts")


```

